Analysis

### 2022-06-04 Logos.Rmd
Logo cloud for CIS-BP human

```{r}
# Modify input dir
dir_data1 <- file.path(dir_data, "Homo_sapiens_2022_05_20_3_16_pm/pwms_CTCF_motifs")
files <- list.files(dir_data1, pattern = "*.txt", full.names = TRUE)
# Output file
fileNameOut1 <- file.path(dir_data1, "CIS-BP_2.00_Homo_sapiens.meme")
fileNameOut2 <- file.path(dir_data1, "CIS-BP_2.00_Homo_sapiens.svg")
# Standard header
write_lines("MEME version 5.4.1 (Tue Mar 1 19:18:48 2022 -0800)", file = fileNameOut1)
write_lines("ALPHABET= ACGT", file = fileNameOut1, append = TRUE)
write_lines("strands: + -", file = fileNameOut1, append = TRUE)
write_lines("Background letter frequencies (from uniform background):", file = fileNameOut1, append = TRUE)
write_lines("A 0.25000 C 0.25000 G 0.25000 T 0.25000 ", file = fileNameOut1, append = TRUE)
# Store number of bases
nsites <- c()
for(i in 1:length(files)) {
  file=files[i]
  # Make motif name
  motif_name <- basename(file) %>% str_remove(., ".txt")
  mtx <- read_tsv(file, col_types = c("ccccc"))
  # If non-empty, proceed
  if (nrow(mtx) > 0) {
    # Collect number of bases
    nsites <- c(nsites, nrow(mtx))
    # Motif name
    write_lines(paste0("MOTIF ", motif_name), file = fileNameOut1, append = TRUE)
    # Header line
    header_line <- paste0("letter-probability matrix: alength= 4 w= ", nrow(mtx), " nsites= 20 E= 0")
    write_lines(header_line, file = fileNameOut1, append = TRUE)
    # Make first numerical column empty
    mtx$Pos <- ""
    # Write actual data
    write_tsv(mtx, file = fileNameOut1, col_names = FALSE, append = TRUE)
    # Separate from the next motif
    write_lines("", file = fileNameOut1, append = TRUE)
  }
}
print("Number of motifs")
print(length(nsites))
print("Summary of motif widths")
summary(nsites)
# Test if can be imported
motifs_all <- importMatrix(fileNameOut1)
if (!file.exists(fileNameOut2)) {
  dev.off()
  svg(fileNameOut2, width = 8, height = 78)
  motifStack(motifs_all, layout="tree") 
  dev.off()
}
```

@@

```{r}
# Modify input file
fileNameIn1 <- file.path(dir_data, "motifs_CTCF.txt")
# Output file
fileNameOut1 <- file.path(dir_data1, "motifs_CTCF.meme")
fileNameOut2 <- file.path(dir_data1, "motifs_CTCF.svg")
# Standard header
write_lines("MEME version 5.4.1 (Tue Mar 1 19:18:48 2022 -0800)", file = fileNameOut1)
write_lines("ALPHABET= ACGT", file = fileNameOut1, append = TRUE)
write_lines("strands: + -", file = fileNameOut1, append = TRUE)
write_lines("Background letter frequencies (from uniform background):", file = fileNameOut1, append = TRUE)
write_lines("A 0.25000 C 0.25000 G 0.25000 T 0.25000 ", file = fileNameOut1, append = TRUE)
# Store number of bases
nsites <- c()
for(i in 1:length(files)) {
  file=files[i]
  mtx <- read.table(fileNameIn1, sep = " ", fill = TRUE)
  # Make motif name
  motif_name <- strsplit(mtx[1, 1], "\t")[[1]]
  motif_id <- motif_name[1] %>% str_remove(">")
  motif_alt <- motif_name[length(motif_name)]
  # If non-empty, proceed
  if (nrow(mtx) > 0) {
    # Collect number of bases
    nsites <- c(nsites, nrow(mtx))
    # Motif name
    write_lines(paste0("MOTIF ", motif_name), file = fileNameOut1, append = TRUE)
    # Header line
    header_line <- paste0("letter-probability matrix: alength= 4 w= ", nrow(mtx), " nsites= 20 E= 0")
    write_lines(header_line, file = fileNameOut1, append = TRUE)
    # Make first numerical column empty
    mtx$Pos <- ""
    # Write actual data
    write_tsv(mtx, file = fileNameOut1, col_names = FALSE, append = TRUE)
    # Separate from the next motif
    write_lines("", file = fileNameOut1, append = TRUE)
  }
}
print("Number of motifs")
print(length(nsites))
print("Summary of motif widths")
summary(nsites)
# Test if can be imported
motifs_all <- importMatrix(fileNameOut1)
if (!file.exists(fileNameOut2)) {
  dev.off()
  svg(fileNameOut2, width = 8, height = 2.5)
  motifStack(motifs_all, layout="tree") 
  dev.off()
}
```

>>>>

```{r}
# Modify input dir
dir_data1 <- file.path(dir_data, "Homo_sapiens_2022_05_20_3_16_pm/pwms_CTCF_motifs")
files <- list.files(dir_data1, pattern = "*.txt", full.names = TRUE)
# Output file
fileNameOut1 <- file.path(dir_data1, "CIS-BP_2.00_Homo_sapiens.meme")
fileNameOut2 <- file.path(dir_data1, "CIS-BP_2.00_Homo_sapiens.svg")
# Standard header
write_lines("MEME version 5.4.1 (Tue Mar 1 19:18:48 2022 -0800)", file = fileNameOut1)
write_lines("ALPHABET= ACGT", file = fileNameOut1, append = TRUE)
write_lines("strands: + -", file = fileNameOut1, append = TRUE)
write_lines("Background letter frequencies (from uniform background):", file = fileNameOut1, append = TRUE)
write_lines("A 0.25000 C 0.25000 G 0.25000 T 0.25000 ", file = fileNameOut1, append = TRUE)
# Store number of bases
nsites <- c()
for(i in 1:length(files)) {
  file=files[i]
  # Make motif name
  motif_name <- basename(file) %>% str_remove(., ".txt") %>% str_replace(., ".", "_")
  mtx <- read_tsv(file, col_types = c("ccccc"))
  # If non-empty, proceed
  if (nrow(mtx) > 0) {
    # Collect number of bases
    nsites <- c(nsites, nrow(mtx))
    # Motif name
    write_lines(paste0("MOTIF ", motif_name), file = fileNameOut1, append = TRUE)
    # Header line
    header_line <- paste0("letter-probability matrix: alength= 4 w= ", nrow(mtx), " nsites= 20 E= 0")
    write_lines(header_line, file = fileNameOut1, append = TRUE)
    # Make first numerical column empty
    mtx$Pos <- ""
    # Write actual data
    write_tsv(mtx, file = fileNameOut1, col_names = FALSE, append = TRUE)
    # Separate from the next motif
    write_lines("", file = fileNameOut1, append = TRUE)
  }
}
print("Number of motifs")
print(length(nsites))
print("Summary of motif widths")
summary(nsites)
# Test if can be imported
motifs_all <- importMatrix(fileNameOut1)
if (!file.exists(fileNameOut2)) {
  dev.off()
  svg(fileNameOut2, width = 8, height = 4)
  # motifStack(motifs_all, layout="tree") 
  ## use matalign to calculate the distances of motifs
  hc <- clusterMotifs(motifs_all)
  ## convert the hclust to phylog object
  library(ade4)
  phylog1 <- ade4::hclust2phylog(hc)
  ## reorder the motifs_all by the order of hclust
  leaves <- names(phylog1$leaves)
  motifs_all <- motifs_all[leaves]
  ## extract the motif signatures
  motifSig <- motifSignature(motifs_all, phylog1, cutoffPval=0.0001, min.freq=1)
  ## draw the motifs with a tag-cloud style.
  motifCloud(motifSig, scale=c(1, 0.55), rot.per = 0,
             layout="rectangles", 
             freq.col = "black")
  dev.off()
}
```

@@

```{r eval=FALSE}
# Modify input file
fileNameIn1 <- file.path(dir_data, "motifs_CTCF.txt")
# Output file
fileNameOut1 <- file.path(dir_data1, "motifs_CTCF.meme")
fileNameOut2 <- file.path(dir_data1, "motifs_CTCF.svg")
# Standard header
write_lines("MEME version 5.4.1 (Tue Mar 1 19:18:48 2022 -0800)", file = fileNameOut1)
write_lines("ALPHABET= ACGT", file = fileNameOut1, append = TRUE)
write_lines("strands: + -", file = fileNameOut1, append = TRUE)
write_lines("Background letter frequencies (from uniform background):", file = fileNameOut1, append = TRUE)
write_lines("A 0.25000 C 0.25000 G 0.25000 T 0.25000 ", file = fileNameOut1, append = TRUE)
# Store number of bases
nsites <- c()
for(i in 1:length(files)) {
  file=files[i]
  mtx <- read.table(fileNameIn1, sep = " ", fill = TRUE)
  # Make motif name
  motif_name <- strsplit(mtx[1, 1], "\t")[[1]]
  motif_id <- motif_name[1] %>% str_remove(">")
  motif_alt <- motif_name[length(motif_name)]
  # If non-empty, proceed
  if (nrow(mtx) > 0) {
    # Collect number of bases
    nsites <- c(nsites, nrow(mtx))
    # Motif name
    write_lines(paste0("MOTIF ", motif_name), file = fileNameOut1, append = TRUE)
    # Header line
    header_line <- paste0("letter-probability matrix: alength= 4 w= ", nrow(mtx), " nsites= 20 E= 0")
    write_lines(header_line, file = fileNameOut1, append = TRUE)
    # Make first numerical column empty
    mtx$Pos <- ""
    # Write actual data
    write_tsv(mtx, file = fileNameOut1, col_names = FALSE, append = TRUE)
    # Separate from the next motif
    write_lines("", file = fileNameOut1, append = TRUE)
  }
}
print("Number of motifs")
print(length(nsites))
print("Summary of motif widths")
summary(nsites)
# Test if can be imported
motifs_all <- importMatrix(fileNameOut1)
if (!file.exists(fileNameOut2)) {
  dev.off()
  svg(fileNameOut2, width = 8, height = 2.5)
  motifStack(motifs_all, layout="tree") 
  dev.off()
}
```
